---
title: "Properties of exponential distribution"
date: "6/25/2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- This app describes the properties of an exponential
distribution.
- The user inputs the rate parameter $\lambda$ and the app draws samples from the exponential distribution and plots the histogram.
- The app also calculates the distribution of the mean of samples drawn from the exponential distribution
- Note: This is different from the taking a large number of draws from the exponential distribution (as shown in the plot in the next slide which looks exponential).


## Exponential distribution

A random variable is said to follow an exponential distribution with rate $\lambda$ is the associated probablity density is 
$$P(x) = \lambda \exp(-\lambda x)$$
and it is defined for $x>0$. It describes the times between events in a Poisson process i.e., a process in which the events occur continuously and independently at a constant average rate $\lambda$.

If $X$ is a random variable with this density, then the expected values of the mean and standard deviation are both
$1/\lambda$.


## Drawing samples

A histogram of 1000 draws from this distribution is given below:

```{r fig.height=4}
lambda <- 0.2
mns0 <- rexp(1000, rate = lambda)
hist(mns0, breaks = 50,
     xlab = "x", ylab = "Frequency",
     main = "")
```


## Sample average of the mean
Now let's sample $n = 40$ random variables from this distribution, 
find its average and repeat the simulation 1000 times. The sample average of the mean and its sample variance can be calculated. 

```{r fig.height=4}
n <- 40
lambda <- 0.2
avg_theory <- 1/lambda

mns <- NULL
for (i in 1 : 1000) mns <- c(mns, mean(rexp(n, rate = lambda)))
avg_sim <- mean(mns)

hist(mns, breaks = 50, density = 10, probability = TRUE,
     xlab = "Mean", ylab = "Density", 
     main = "")
rug(mns, col = "green")
abline(v = avg_sim, col = "red")
abline(v = avg_theory, col = "blue")
legend("topright", 
       legend=c(paste("Sample Mean = ", round(avg_sim,3)),
                paste("Theoretical Mean = ", avg_theory)),
       col=c("red", "blue"), lty=1, cex=0.8)
lines(density(mns), col="black", lwd=2)
```

